\chapter*{Acknowledgments}
\label{ch:acks}

During the past 9 years at Columbia University, I have met inspiring advisors
and colleagues who defined who I am as a researcher. My deepest gratitude goes
to my main advisor, Jason Nieh, for believing in me and making me reach my full
potential. He never gave up on me, even when he had all the reasons to. I am
grateful to my first secondary advisor, Roxana Geambasu, for teaching me
distributed systems design in a rigorous and practical manner. She offered
countless hours of guidance both on a professional level and personal level. I
could always count on her when I needed a listening ear. I am grateful to my
second secondary advisor, Junfeng Yang, for elevating the conversations. His
knowledge and expertise in systems were invaluable in developing my work. I am
grateful to my mentor, Oren Laadan, for providing guidance when I started my
research. His ability to express complex concepts with such ease is fascinating.
I wish I could be as articulated and eloquent as he is. I am grateful to my
friend and colleague, Sid Nair, for making me a better person by touching me
with kindness, softness, and humility. Interacting with Sid has always been a
wonderful, even during disagreements. I am grateful to Kareem Kouddous for
allowing me to implement my research prototype in a large production system he
was responsible for. He brought back the self-esteem I had once lost. I am
grateful to Jeremy Andrus for consistently giving me a hand when I struggled
getting things done. I have tremendous respect for his sense of discipline and
commitment. I am grateful to Christoffer Dall for being a source of inspiration.
He became lead developer of the ARM/KVM subsystem of the Linux kernel thanks to
his social, communication, and technical skills. I want to thank all the
Columbia University staff that made a difference to me. To name a few, Jessica
Rosa, Remiko Moss, Daisy Nguyen, and Cindy Walters. I also want to thank my
friends Valentin Burlacu, Deirdre Campbell, Mathias Lecuyer, Quentin Decock,
Aaron Allon, and Laure Theroude for providing their support for many years.
Finally, I'd like to acknowledge myself for finishing what I thought impossible
to complete.

This work was supported in part by an IBM Faculty Award, a Google Research
Award, NSF grants CNS-1018355, CNS-0905246, CCF-1162021, NCNS-1162447,
CNS-1422909, CNS-1351089, CNS-1117805, CNS-1054906 (CAREER), CNS-1012633,
CNS-0914845, CNS-0905246, AFRL FA8650-10-C-7024, AFRL FA8750-10-2-0253, AFOSR MURI
FA9550-07-1-0527, and DARPA FA8650-11-C-7190.
